---
output: github_document
bibliography: bibliography.bib  
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rOPTRAM

<!-- badges: start -->
[![Project Status: WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

`rOPTRAM` implements The OPtical TRapezoid Model (OPTRAM) 
  to derive soil moisture based on the linear relation
  between a vegetation index, i.e. NDVI, and Land Surface Temperature (LST).
  
  The Short Wave Infra-red (SWIR) band is used as a proxy for LST. The SWIR band is transformed to Swir Transformed Reflectance (STR).
  
  A scatterplot of NDVI vs. STR is used to produce wet and dry linear regression lines, and the slope/intercept coefficents of these lines comprise the trapezoid. These coefficients are then used on a new satellite image to determine soil moisture.

  See:  @sadeghi_optical_2017, @burdun_satellite_2020, @ambrosone_retrieving_2020


## Installation

`rOPTRAM` resides on gitlab.
You can install the development version of rOPTRAM like so:

``` {r install, eval=FALSE}
# Install remotes package
if (! require("remotes")) {
  install.packages("remotes")
}
remotes::install_gitlab("rsl-bidr/roptram")
```
#### Suggested prerequisites

This package works with the R package `sen2r`.
To install that package you will need some system dependencies. 
On Debian and derivitives please install in advance:
 
  - jq libv8-dev libjq-dev
  - libprotobuf-dev protobuf-compiler

Install the google cloud CLI to allow acquiring images directly from Google.
For installation instructions see:
https://cloud.google.com/sdk/docs/install
 
## Example

This is a basic example which shows how to:

  - retrieve Sentinel 2 imagery for a specific area of interest
  - covering a date range
  - preprocess the imagery to obtain a vegetation index and STR band
  - use these to derive coefficients of slope and intercept for the OPTRAM trapezoid

```{r example, eval=FALSE}
library(rOPTRAM)
# basic example
geojson_file <- system.file("extdata", "migda_9.geojson", package="rOPTRAM")
coeffs <- rOPTRAM::optram(
    aoi = geojson_file,
    from_date = "2021-01-01", to_date = "2021-03-30",
    max_cloud = 10,
    vi = "NDVI")
coeffs
```

## Note

In order to download Sentinel 2 images, the `sen2r` package is used:

@ranghetti_sen2r_2020

That package must be installed in advance in order to run the `optram()` wrapper function.

If, on the other hand, Sentinel 2 imagery has been downloaded in advance, then `sen2r` is not necessary.
Instead, the following workflow can be used:
  - call `optram_safe()` to prepare NDVI and STR rasters
  - call `optram_calculate_str()` to calculate SWIR Transform
  - call `optram_ndvi_str()` to build a data.frame of pixel values
  - call `optram_wetdry_coefficients()` to derive slope and intercept of the trapezoid. 

## Meta

  - Please report any issues on [gitlab](https://gitlab.com/rsl-bidr/roptram/-/issues)
  - Anyone interested in collaborating is invited to "sign up" by contacting the maintainers.
  - This package is released with a [Contributor Code of Conduct](https://github.com/ropensci/.github/blob/master/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
